---
title: "A03_G46086231"
author: "Min Yang"
date: "9/14/2018"
output: html_document
---
```{r}
library(dplyr)
library(ggplot2)
```

```{r}
kf <- read.csv("~/Desktop/kicksfootball.csv") #input data
kf$g <- ifelse(kf$goal=="Y", 1, 0)  #convert Y and N to 1, 0
```

##Part A
```{r}
sum(kf$g)/length(kf$g) #overall probability of hit field goal 
```
```{r}
match <-subset(kf$g, kf$practiceormatch=="M") #subset goal of matching game 
prac <-subset(kf$g, kf$practiceormatch=="P") #subset goal of practice 
probmatch <- sum(match)/length(match) #calculate probability of hit goal in match
probprac <- sum(prac)/length(prac) #calculate probability of hit goal in practice
probmatch
probprac
```

The overall probabtily of Melvin to score a goal is 0.779,the probability of him hit a goal during practice session is 0.788, and the probability of him hit a goal during matches is 0.701.  

```{r}
gp<- plot(g~yards,data=kf,
     xlab="yards",
     ylab="Probability of a goal",
     main="Overall Performance")
log = glm(g ~ yards,
            data=kf,
            family=binomial)
curve(predict(log,
              data.frame(yards=x),
              type="resp"),
     add=TRUE)
```

From the plot, we can see the trend that the probability of a goal decrease as yards increase, so I would predict Melvin has a high probability around 90% if he kick at 20 yards, and a decreasing probability of hit the goal as yards increase, and around 30% of success at 60 yards.  

##Part B
Melvin’s attempts in matches is different from his performance in practice, he generally has a better performace in practice than match.

###a) 

### i. Practice
```{r}
prac <- kf %>% select(practiceormatch, g, yards)%>% filter(practiceormatch == "P")
prac.log = glm(g ~ yards,
            data=prac,
            family=binomial)
summary(prac.log)
```

logistic function for practice: y = 4.49 - 0.082x


### ii. Match
```{r}
match <- kf %>% select(practiceormatch, g, yards)%>% filter(practiceormatch == "M")
match.log = glm(g ~ yards,
            data=match,
            family=binomial)
summary(match.log)
```

logistic function for match: y = 4.26 - 0.088x

##b) probability in practice
### 20 yards
```{r}
inp <- c(20)
newdata = data.frame(yards=inp)
predict(prac.log, newdata, type="response")
```

In practice, the probability of Melvin scoring a goal from 20 yards is 0.9451.

### 40 yards
```{r}
inp <- c(40)
newdata = data.frame(yards=inp)
predict(prac.log, newdata, type="response")
```

In practice, the probability of Melvin scoring a goal from 40 yards is 0.7683.

### 60 yards
```{r}
inp <- c(60)
newdata = data.frame(yards=inp)
predict(prac.log, newdata, type="response")
```

In practice, the probability of Melvin scoring a goal from 60 yards is 0.3897.

##c) probability in match 
### 20 yards
```{r}
inp <- c(20)
newdata = data.frame(yards=inp)
predict(match.log, newdata, type="response")
```

In match, the probability of Melvin scoring a goal from 20 yards is 0.9236.

### 40 yards
```{r}
inp <- c(40)
newdata = data.frame(yards=inp)
predict(match.log, newdata, type="response")
```

In match, the probability of Melvin scoring a goal from 40 yards is 0.6737.

### 60 yards
```{r}
inp <- c(60)
newdata = data.frame(yards=inp)
predict(match.log, newdata, type="response")
```

In match, the probability of Melvin scoring a goal from 60 yards is 0.2606.

```{r}
matchplot <- plot(g~yards,data=match,
     xlab="yards",
     ylab="Probability of a goal",
     main="Match")
curve(predict(match.log,
              data.frame(yards=x),
              type="resp"),
     add=TRUE)
```

```{r}
plot(g~yards,data=prac,
     xlab="yards",
     ylab="Probability of a goal",
     main="Practice")
curve(predict(prac.log,
              data.frame(yards=x),
              type="resp"),
     add=TRUE)
```

##Part C
```{r}
  g<- ggplot(kf) + aes(x=yards,y=g, color= practiceormatch)+labs(title="Melvin Performance with Comparison in Match and Practice",x="Yards",y="Probability of a goal")+
  geom_point(data = match, size = 2)+
  geom_point(data = prac, size = 1)
 g+ stat_smooth(method="glm", method.args = list(family="binomial"), data = match, formula = y ~ x, se=FALSE)+
  stat_smooth(method="glm", method.args = list(family="binomial"), data = prac, formula = y ~ x, se=FALSE)

```



